---
title: "`r params$author_name` — PubliGraphics"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  html_document:
    theme: null
    highlight: null
    css: ["publigraphics_base.css"]
    self_contained: true
    toc: false
    fig_caption: true
    df_print: kable
  pagedown::html_paged:
    css: ["publigraphics_base.css", "publigraphics_print.css"]
    self_contained: true
params:
  author_name: ""
  affiliation: ""
  theme_color: "#1B4F72"
  language: "fr"
  cover_html: ""
  data: NULL
  summary_table: NULL
  narratives: NULL
  viz_dir: ""
  viz_paths: NULL
  stats_list: NULL
  sii_result: NULL
  sii_card_html: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 9,
  fig.height = 6,
  dpi = 150,
  out.width = "100%"
)

# --- Bilingual header helper ------------------------------------------------
h <- function(fr, en) {
  if (identical(params$language, "fr")) fr else en
}

# --- Unpack parameters -------------------------------------------------------
data       <- params$data
color      <- params$theme_color
viz_paths  <- params$viz_paths
stats_list <- params$stats_list
sii_result <- params$sii_result

# --- Helper: include a viz image if it exists --------------------------------
include_viz <- function(filename) {
  if (!is.null(viz_paths) && filename %in% names(viz_paths)) {
    path <- viz_paths[[filename]]
    if (file.exists(path)) {
      knitr::include_graphics(path)
    }
  }
}
```

```{css, echo=FALSE}
:root {
  --pg-theme-color: `r params$theme_color`;
}
```

<!-- ======================================================================= -->
<!-- SECTION 1 : COVER PAGE                                                  -->
<!-- ======================================================================= -->

<div class="pg-cover-page">
`r params$cover_html`
</div>

<div class="pg-page-break"></div>

<!-- ======================================================================= -->
<!-- SECTION 2 : OVERVIEW                                                    -->
<!-- ======================================================================= -->

## `r h("Vue d'ensemble", "Overview")` {.pg-section}

<div class="pg-stats-grid">
```{r stats-cards, results='asis'}
if (!is.null(stats_list)) {
  labels <- if (identical(params$language, "fr")) {
    c("Articles", "Ouvrages", "S\u00e9minaires", "Projets", "Th\u00e8ses", "Prix")
  } else {
    c("Articles", "Books", "Seminars", "Projects", "Theses", "Awards")
  }
  values <- c(
    stats_list$Articles, stats_list$Books, stats_list$Seminars,
    stats_list$Projects, stats_list$Theses, stats_list$Awards
  )
  for (i in seq_along(labels)) {
    cat(sprintf(
      '<div class="pg-stat-card" style="border-top: 3px solid %s;">
         <div class="pg-stat-value">%s</div>
         <div class="pg-stat-label">%s</div>
       </div>\n',
      color, values[i], labels[i]
    ))
  }
}
```
</div>

```{r wordcloud-overview, fig.cap=h("Mots-cl\u00e9s des articles", "Article keywords")}
include_viz("wordcloud_articles.png")
```

```{r timeline-overview, fig.cap=h("Chronologie des publications", "Publication timeline")}
include_viz("timeline_articles.png")
```

<div class="pg-page-break"></div>

<!-- ======================================================================= -->
<!-- SECTION 3 : SCIENTIFIC INFLUENCE INDEX (SII)                            -->
<!-- ======================================================================= -->

## `r h("Indice d'Influence Scientifique", "Scientific Influence Index")` {.pg-section}

```{r sii-card, results='asis'}
if (nchar(params$sii_card_html) > 0) {
  cat(params$sii_card_html)
}
```

```{r sii-radar, fig.cap=h("Radar multidimensionnel SII", "SII multi-dimensional radar")}
include_viz("sii_radar.png")
```

```{r sii-evolution, fig.cap=h("\u00c9volution temporelle du SII", "SII temporal evolution")}
include_viz("sii_evolution.png")
```

<div class="pg-page-break"></div>

<!-- ======================================================================= -->
<!-- SECTION 4 : ARTICLES                                                    -->
<!-- ======================================================================= -->

## `r h("Articles", "Articles")` {.pg-section}

### `r h("Top 5 des articles", "Top 5 Articles")`

```{r top-articles, results='asis'}
if (!is.null(data)) {
  articles <- data[data$type_classified == "article", ]
  if (nrow(articles) > 0) {
    # Sort by cited_by descending, then by year descending
    if ("cited_by" %in% names(articles)) {
      articles <- articles[order(-articles$cited_by, -articles$year), ]
    } else {
      articles <- articles[order(-articles$year), ]
    }
    top5 <- head(articles, 5)
    for (i in seq_len(nrow(top5))) {
      row <- top5[i, ]
      year_badge <- sprintf(
        '<span class="pg-badge" style="background-color: %s; color: #fff;">%s</span>',
        color, row$year
      )
      cited_badge <- ""
      if ("cited_by" %in% names(row) && !is.na(row$cited_by) && row$cited_by > 0) {
        cited_badge <- sprintf(
          '<span class="pg-badge-outline">%s citations</span>',
          row$cited_by
        )
      }
      journal_text <- if ("journal" %in% names(row) && !is.na(row$journal) &&
                          nchar(as.character(row$journal)) > 0) {
        sprintf('<em>%s</em>', row$journal)
      } else {
        ""
      }
      cat(sprintf(
        '<div class="pg-article-card" style="border-left-color: %s;">
           <h4>%s %s %s</h4>
           <p>%s</p>
           <p>%s</p>
         </div>\n',
        color, row$title, year_badge, cited_badge, row$authors, journal_text
      ))
    }
  }
}
```

<div class="pg-page-break"></div>

<!-- ======================================================================= -->
<!-- SECTION 5 : BOOKS                                                       -->
<!-- ======================================================================= -->

## `r h("Ouvrages", "Books")` {.pg-section}

### `r h("Galerie des ouvrages", "Book Gallery")`

```{r book-gallery, fig.cap=h("Galerie des ouvrages", "Book gallery")}
include_viz("gallery_books.png")
```

### `r h("R\u00e9seau de co-autorat", "Co-authorship Network")`

```{r coauthorship-network, fig.cap=h("R\u00e9seau de co-autorat", "Co-authorship network")}
include_viz("network_coauthors.png")
```

<div class="pg-page-break"></div>

<!-- ======================================================================= -->
<!-- SECTION 6 : SEMINARS                                                    -->
<!-- ======================================================================= -->

## `r h("S\u00e9minaires & Conf\u00e9rences", "Seminars & Conferences")` {.pg-section}

### `r h("Carte des interventions", "Presentation Map")`

```{r seminar-map, fig.cap=h("Localisation des s\u00e9minaires", "Seminar locations")}
include_viz("map_seminars_static.png")
```

### `r h("R\u00e9seau th\u00e9matique", "Thematic Network")`

```{r thematic-network, fig.cap=h("R\u00e9seau th\u00e9matique des s\u00e9minaires", "Seminar thematic network")}
include_viz("network_seminars.png")
```

<div class="pg-page-break"></div>

<!-- ======================================================================= -->
<!-- SECTION 7 : PROJECTS & THESES                                           -->
<!-- ======================================================================= -->

## `r h("Projets & Th\u00e8ses", "Projects & Theses")` {.pg-section}

### `r h("Diagramme de Gantt des projets", "Project Gantt Chart")`

```{r gantt-chart, fig.cap=h("Gantt des projets de recherche", "Research project Gantt chart")}
include_viz("gantt_projects.png")
```

### `r h("Chronologie des th\u00e8ses", "Thesis Timeline")`

```{r thesis-timeline, fig.cap=h("Chronologie des directions de th\u00e8se", "Thesis supervision timeline")}
include_viz("timeline_theses.png")
```

<div class="pg-page-break"></div>

<!-- ======================================================================= -->
<!-- SECTION 8 : AWARDS & MEDIA                                              -->
<!-- ======================================================================= -->

## `r h("Prix & M\u00e9dias", "Awards & Media")` {.pg-section}

### `r h("Distinctions", "Awards")`

```{r awards-infographic, fig.cap=h("Infographie des distinctions", "Awards infographic")}
include_viz("infographic_awards.png")
```

### `r h("Synth\u00e8se m\u00e9diatique", "Media Summary")`

```{r media-summary, fig.cap=h("Synth\u00e8se m\u00e9diatique", "Media summary")}
include_viz("media_summary.png")
```

---

<div style="text-align: center; padding: 30px 0; color: #888; font-size: 0.85em;">
`r h("G\u00e9n\u00e9r\u00e9 avec", "Generated with")` <strong>PubliGraphics</strong> `r packageVersion("publigraphics")` — `r format(Sys.time(), '%d %B %Y')`
</div>
