---
title: "Advanced customization"
author: "Paul Wambo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced customization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette covers advanced customisation options for publigraphics. If
you are new to the package, start with `vignette("introduction")`.

## Custom themes with `pg_theme()` and `pg_palette()`

Every visualisation produced by publigraphics uses the internal ggplot2
theme `pg_theme()` and the colour palette generator `pg_palette()`. Both
are exported and can be used directly in your own plots or overridden
globally.

### `pg_theme()`

`pg_theme()` returns a publication-ready ggplot2 theme built on top of
`theme_minimal()`. It loads the Lato font family via `sysfonts` /
`showtext` and applies clean defaults for titles, gridlines, and margins.

```{r theme-basic}
library(publigraphics)
library(ggplot2)

# Default theme
ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  pg_theme()

# Custom base size and accent colour
ggplot(mtcars, aes(wt, mpg)) +
  geom_point(colour = "#E74C3C") +
  pg_theme(base_size = 14, base_color = "#E74C3C")
```

**Parameters**:

| Argument | Description | Default |
|:------------|:--------------------------------------|:------------|
| `base_size` | Base font size in points | `12` |
| `base_color`| Hex colour for titles and accents | `"#1B4F72"` |

To set `pg_theme()` as the default for all subsequent plots in a session:

```{r theme-set}
theme_set(pg_theme(base_color = "#2E86C1"))
```

### `pg_palette()`

`pg_palette()` generates a vector of `n` hex colour codes derived from a
base colour using the `colorspace` package. Three palette types are
available:

- **sequential**: a gradient from dark to light (good for continuous data)
- **diverging**: two hues spreading from a neutral centre
- **qualitative**: distinct hues for categorical data

```{r palette-examples}
# 5 sequential colours from a deep blue base
pg_palette("#1B4F72", n = 5, type = "sequential")

# 4 qualitative colours from a red base
pg_palette("#E74C3C", n = 4, type = "qualitative")

# 7 diverging colours
pg_palette("#27AE60", n = 7, type = "diverging")
```

You can use these palettes with `scale_colour_manual()` or
`scale_fill_manual()` in any ggplot:

```{r palette-use}
pal <- pg_palette("#1B4F72", n = 3, type = "qualitative")

ggplot(iris, aes(Sepal.Length, Sepal.Width, colour = Species)) +
  geom_point(size = 2) +
  scale_colour_manual(values = pal) +
  pg_theme()
```

## Using your own data

The demo data shipped with publigraphics (Esther Duflo's publications)
is useful for testing, but you will typically want to use your own
bibliographic records.

### BibTeX from Zotero or Mendeley

Export your library as a `.bib` file from Zotero (File > Export Library >
BibTeX) or Mendeley (Tools > Export > BibTeX). Then pass the path to
`pg_read_bib()`:
```{r own-bib}
my_refs <- pg_read_bib("path/to/my_library.bib")
```

### CSV or XLSX files

If your data is in a spreadsheet, export it as CSV. Column names are
matched case-insensitively to the publigraphics schema. The most important
columns are:

| Column | Required? | Description |
|:-----------------|:----------|:--------------------------------------|
| `title` | Yes | Title of the publication |
| `year` | Yes | Publication year (integer) |
| `authors` | Yes | Author names (semicolon-separated) |
| `type_raw` | Recommended | BibTeX entry type or free text |
| `journal` | Optional | Journal name or venue |
| `abstract` | Optional | Abstract text (used for word clouds) |
| `keywords` | Optional | Comma-separated keywords |
| `doi` | Optional | Digital Object Identifier |

```{r own-csv}
my_refs <- pg_read_bib("path/to/my_publications.csv")
```

### Supplementary CSV for non-bibliographic outputs

For items not captured in standard bibliographic databases (seminars,
funded projects, supervised theses, awards, consulting), create a CSV with
at least the columns `type`, `title`, and `year`:

```{r own-extra}
my_extras <- pg_read_extra("path/to/my_extras.csv")
```

Expected columns include: `type` (e.g. "seminar", "project", "award"),
`title`, `year`, `city`, `country`, `institution`, `note`,
`funding_amount`, and `funding_source`. Missing columns default to `NA`.

### Full pipeline with your own data

```{r own-pipeline}
library(publigraphics)

refs   <- pg_read_bib("my_library.bib")
extras <- pg_read_extra("my_extras.csv")

generate_publigraphics(
  author_name    = "Lastname, Firstname",
  bib_file       = "my_library.bib",
  extra_data     = "my_extras.csv",
  photo          = "my_photo.jpg",
  affiliation    = "My University",
  orcid          = "0000-0000-0000-0000",
  website        = "https://mywebsite.com",
  theme_color    = "#2E86C1",
  language       = "en",
  output_formats = c("html", "pdf")
)
```

## AI narrative generation

publigraphics can generate structured narrative summaries for each
production using the Anthropic Claude API. These narratives are
automatically included in the notebook when an API key is available.

### Setting up your API key

Set the `ANTHROPIC_API_KEY` environment variable before calling the
pipeline. The recommended approach is to add it to your `.Renviron` file:

```{r api-key-renviron}
# In your .Renviron file (edit with usethis::edit_r_environ()):
# ANTHROPIC_API_KEY=sk-ant-api03-xxxxxxxxxxxx
```

Alternatively, pass the key directly:

```{r api-key-direct}
generate_publigraphics(
  author_name  = "Lastname, Firstname",
  bib_file     = "my_library.bib",
  api_key_claude = "sk-ant-api03-xxxxxxxxxxxx",
  language     = "fr"
)
```

### Calling `pg_run_all_narratives()` directly

If you want finer control over narrative generation, you can call the
function directly on a classified dataset:

```{r narratives-direct}
library(publigraphics)

data <- pg_read_bib("my_library.bib") |>
  pg_merge_inputs() |>
  pg_classify()

enriched <- pg_run_all_narratives(
  data      = data,
  api_key   = Sys.getenv("ANTHROPIC_API_KEY"),
  lang      = "en",
  max_items = 10
)
```

The function adds six columns to the tibble:

| Column | Description |
|:--------------------------------|:------------------------------------------|
| `narrative_problematique` | Research question or central theme |
| `narrative_pertinence` | Importance and context |
| `narrative_resultat` | Main findings or contribution |
| `narrative_question_ouverte` | Open question for further reflection |
| `narrative_lang` | Language code used for generation |
| `narrative_cached` | Whether the result was loaded from cache |

### Caching

To avoid redundant API calls on re-runs, pass a `cache_dir`:

```{r narratives-cache}
enriched <- pg_run_all_narratives(
  data      = data,
  api_key   = Sys.getenv("ANTHROPIC_API_KEY"),
  lang      = "fr",
  cache_dir = "narrative_cache"
)
```

Each result is keyed by a hash of `pg_id`, `abstract`, and `lang`. On
subsequent runs, cached results are loaded from `.rds` files instead of
calling the API.

### Controlling the scope

Use the `types_to_process` and `max_items` arguments to limit which
productions receive narratives:

```{r narratives-scope}
enriched <- pg_run_all_narratives(
  data             = data,
  api_key          = Sys.getenv("ANTHROPIC_API_KEY"),
  lang             = "en",
  types_to_process = c("article", "book"),
  max_items        = 5
)
```

## Customising the cover page

The cover page is built by `pg_build_cover()`, which generates a
self-contained HTML string with a two-column CSS Grid layout: a coloured
left accent band and a right content area with the researcher's photo,
name, affiliation, social links, a statistics grid, and an optional QR
code.

### Basic usage

```{r cover-basic}
cover_html <- pg_build_cover(
  author_name = "Marie Dupont",
  affiliation = "Universite de Paris",
  photo_path  = "photo.jpg",
  orcid       = "0000-0001-2345-6789",
  linkedin    = "https://linkedin.com/in/mariedupont",
  website     = "https://mariedupont.fr",
  stats       = list(
    Articles = 24, Books = 3, Seminars = 15,
    Theses = 8, Awards = 2, Projects = 5
  ),
  theme_color = "#1B4F72",
  lang        = "fr"
)

# Preview in browser
htmltools::browsable(htmltools::HTML(cover_html))
```

### Cover parameters

| Parameter | Description |
|:--------------|:------------------------------------------------------|
| `author_name` | Full researcher name (displayed prominently) |
| `affiliation` | Institution name |
| `photo_path` | Path to JPEG/PNG photo (or `NULL` for initials) |
| `orcid` | ORCID identifier (displayed as clickable link) |
| `linkedin` | LinkedIn profile URL |
| `website` | Personal website URL (generates a QR code) |
| `stats` | Named list with 6 elements: Articles, Books, Seminars, Theses, Awards, Projects |
| `theme_color` | Hex colour for the accent band and headings |
| `lang` | `"fr"` or `"en"` for label language |

### Photo handling

When `photo_path` is provided, the image is embedded as a base64 data URI
to keep the output self-contained. When `NULL`, an initials placeholder
(a coloured circle with the researcher's initials) is generated
automatically from `author_name`.

### Changing the accent colour

The left accent band uses a gradient from `theme_color` to a darker shade
computed via `colorspace::darken()`. Changing `theme_color` affects the
entire visual identity of both the cover and the notebook:

```{r cover-colour}
# A warm red theme
cover_red <- pg_build_cover(
  author_name = "Marie Dupont",
  affiliation = "Universite de Paris",
  stats       = list(
    Articles = 24, Books = 3, Seminars = 15,
    Theses = 8, Awards = 2, Projects = 5
  ),
  theme_color = "#C0392B",
  lang        = "en"
)
```

## Custom CSS for the notebook

The rendered notebook is an HTML document. You can inject custom CSS to
override styling by creating a CSS file and referencing it in your
workflow. For example, to change heading fonts or the page background:
```{r custom-css}
# Create a custom CSS file
css_content <- "
/* Custom publigraphics overrides */
body {
  font-family: 'Georgia', serif;
  background-color: #FAFAFA;
}

h1, h2, h3 {
  font-family: 'Montserrat', sans-serif;
  color: #2C3E50;
}

.pg-stat-card {
  border-radius: 12px;
  box-shadow: 0 2px 8px rgba(0, 0, 0, 0.08);
}

/* Narrower page margins for print */
@media print {
  body {
    margin: 1cm;
  }
}
"
writeLines(css_content, "my_custom_styles.css")
```

You can include this CSS file in the R Markdown output by adding it to
the YAML header of a custom template or by injecting it via
`htmltools::tags$style()` in a setup chunk.

## Combining multiple BibTeX files

If your publications are split across several BibTeX files (e.g., one
from Zotero and one from Google Scholar), you can read them separately and
merge:

```{r multi-bib}
refs_zotero <- pg_read_bib("zotero_export.bib")
refs_scholar <- pg_read_bib("scholar_export.bib")

# Combine and deduplicate
all_refs <- pg_merge_inputs(
  bib_data = dplyr::bind_rows(refs_zotero, refs_scholar)
)
```

`pg_merge_inputs()` automatically removes duplicates based on normalised
title comparison. The first occurrence (from the first data frame) is
retained.

## Controlling the visual theme globally

For a consistent look across an entire analysis session, set the theme and
palette globally:

```{r global-theme}
library(publigraphics)
library(ggplot2)

# Set the publigraphics theme as default
theme_set(pg_theme(base_size = 13, base_color = "#8E44AD"))

# All subsequent ggplots use this theme automatically
ggplot(mtcars, aes(hp, mpg, colour = factor(cyl))) +
  geom_point(size = 3) +
  scale_colour_manual(
    values = pg_palette("#8E44AD", n = 3, type = "qualitative")
  ) +
  labs(title = "Example with custom purple theme")
```

## Summary of key functions

| Function | Purpose |
|:-------------------------------|:----------------------------------------------|
| `pg_read_bib()` | Read BibTeX / CSV / XLSX bibliography files |
| `pg_read_extra()` | Read supplementary CSV |
| `pg_merge_inputs()` | Merge and deduplicate data |
| `pg_classify()` | Classify productions into 13 types |
| `pg_summary_table()` | Aggregated summary with bilingual labels |
| `pg_theme()` | Publication-ready ggplot2 theme |
| `pg_palette()` | Colour palette generator |
| `pg_build_cover()` | HTML cover page |
| `pg_run_all_narratives()` | AI narrative summaries (Anthropic API) |
| `generate_publigraphics()` | Full pipeline: data to notebook |

For further details on each function, consult the package reference
documentation with `?function_name` or visit the pkgdown site at
<https://Paul1991-ux.github.io/publigraphics>.
