---
title: "Introduction to publigraphics"
author: "Paul Wambo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to publigraphics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

**publigraphics** generates visually impactful, publication-ready notebooks
(HTML and PDF) that summarise the complete scientific trajectory of a
researcher in the social sciences. Starting from standard bibliographic
exports (BibTeX, RIS, or CSV from Zotero, Mendeley, etc.) and an optional
supplementary CSV for non-bibliographic outputs, the package produces:
thematic word clouds with TF-IDF weighting, co-authorship networks,
publication timelines, geographic maps of seminar interventions, Gantt
charts of funded projects, and AI-generated narrative summaries via the
Anthropic Claude API.

**Bilingual support**: all labels, messages, and AI narratives can be
produced in French or English by setting the `lang` (or `language`)
parameter. The default is French (`"fr"`).

## Installation

You can install the development version of publigraphics from GitHub:

```{r install}
# install.packages("remotes")
remotes::install_github("Paul1991-ux/publigraphics")
```

## Quick start

### Step 1: Load the package

```{r load}
library(publigraphics)
```

### Step 2: Read your data

publigraphics ships with demo data based on a selection of publications by
Esther Duflo. Three files are included under `inst/extdata/`:

- `duflo_articles.bib` -- journal articles in BibTeX format
- `duflo_books.bib` -- books and book chapters in BibTeX format
- `duflo_extra.csv` -- supplementary items (seminars, awards, projects)

Use `pg_read_bib()` to import BibTeX (or CSV/XLSX) bibliography files, and
`pg_read_extra()` for the supplementary CSV:

```{r read-data}
# Locate the bundled demo files
bib_articles <- system.file("extdata", "duflo_articles.bib",
                            package = "publigraphics")
bib_books    <- system.file("extdata", "duflo_books.bib",
                            package = "publigraphics")
extra_csv    <- system.file("extdata", "duflo_extra.csv",
                            package = "publigraphics")

# Read bibliographic data
articles <- pg_read_bib(bib_articles)
books    <- pg_read_bib(bib_books)

# Read supplementary (non-bibliographic) data
extras   <- pg_read_extra(extra_csv)
```

Each function returns a standardised tibble with 19 columns (the
**publigraphics schema**): `pg_id`, `type_raw`, `type_classified`, `title`,
`authors`, `year`, `journal_or_venue`, `abstract`, `keywords`, `doi`,
`url`, `isbn`, `city`, `country`, `institution`, `note`, `cited_by`,
`source`, and `date_added`.

### Step 3: Merge inputs

Combine the bibliographic and supplementary data into a single tibble.
Duplicates are detected by case-insensitive title comparison and
automatically removed (the bibliographic entry is preferred because it
typically carries richer metadata):

```{r merge}
# Merge articles and books first
bib_all <- pg_merge_inputs(
  bib_data   = dplyr::bind_rows(articles, books),
  extra_data = extras
)

nrow(bib_all)
```

### Step 4: Classify productions

`pg_classify()` assigns one of 13 canonical types to each row based on the
BibTeX entry type and discriminant fields. Types include `article`, `book`,
`book_chapter`, `seminar`, `conference`, `report`, `thesis_supervised`,
`patent`, `media`, `project`, `award`, `expertise`, and `other`:

```{r classify}
classified <- pg_classify(bib_all)

table(classified$type_classified)
```

### Step 5: Summary table

`pg_summary_table()` aggregates counts, year ranges, and percentages by
production type, with bilingual labels (French and English):

```{r summary}
summary_tbl <- pg_summary_table(classified)
summary_tbl
```

The resulting tibble contains columns `type_classified`, `label_fr`,
`label_en`, `n`, `first_year`, `last_year`, and `pct_total`.

## A first visualisation

publigraphics includes a wide range of visualisation functions. Here is an
example of a publication timeline for articles:

```{r timeline, fig.width = 10, fig.height = 5}
pg_timeline_articles(classified, theme_color = "#1B4F72")
```

And a thematic word cloud built from article titles and abstracts using
TF-IDF weighting:

```{r wordcloud, fig.width = 9, fig.height = 7}
wc <- pg_wordcloud_articles(classified, theme_color = "#1B4F72")
wc$plot
```

All visualisation functions accept a `theme_color` argument (a hex colour
string) that controls the primary accent colour of the plot. You can also
use `pg_theme()` and `pg_palette()` to customise the appearance globally
(see `vignette("advanced_customization")`).

## Generating the full notebook

The main entry point is `generate_publigraphics()`, which orchestrates the
entire pipeline: data import, classification, statistics, AI narrative
summaries, all visualisations, cover page, R Markdown rendering, and
optional PDF conversion.

```{r generate}
generate_publigraphics(
  author_name    = "Duflo, Esther",
  bib_file       = bib_articles,
  extra_data     = extra_csv,
  affiliation    = "Massachusetts Institute of Technology",
  orcid          = "0000-0003-2404-5946",
  website        = "https://economics.mit.edu/people/faculty/esther-duflo",
  theme_color    = "#1B4F72",
  output_formats = c("html", "pdf"),
  language       = "en",
  open_after     = TRUE
)
```

The function returns an invisible list containing the paths to the
generated files (`html_path`, `pdf_path`), the classified data tibble, and
the wall-clock duration of the pipeline.

### Key parameters

| Parameter | Description | Default |
|:-----------------|:-------------------------------------------|:--------------|
| `author_name` | Full name of the researcher | (required) |
| `bib_file` | Path to `.bib`, `.csv`, or `.xlsx` file | (required) |
| `extra_data` | Path to supplementary CSV (or `NULL`) | `NULL` |
| `photo` | Path to a JPEG/PNG photo for the cover | `NULL` |
| `affiliation` | Institutional affiliation | `""` |
| `orcid` | ORCID identifier | `NULL` |
| `api_key_claude` | Anthropic API key for AI narratives | `Sys.getenv("ANTHROPIC_API_KEY")` |
| `theme_color` | Hex colour code for the visual theme | `"#1B4F72"` |
| `language` | `"fr"` or `"en"` | `"fr"` |
| `output_formats` | `"html"`, `"pdf"`, or both | `c("pdf", "html")` |
| `open_after` | Open HTML in browser when done? | `TRUE` |

## Language support

publigraphics is bilingual throughout. Set `language = "fr"` for French or
`language = "en"` for English. This affects:
- all CLI progress messages (displayed in both languages by default),
- the summary table labels (`label_fr` / `label_en`),
- the cover page labels and footer,
- AI narrative prompts when using the Anthropic API.

```{r language-example}
# French notebook (default)
generate_publigraphics(
  author_name = "Duflo, Esther",
  bib_file    = bib_articles,
  language    = "fr"
)

# English notebook
generate_publigraphics(
  author_name = "Duflo, Esther",
  bib_file    = bib_articles,
  language    = "en"
)
```

## What's next?

- **Custom themes and palettes**: see `vignette("advanced_customization")`
- **AI narratives**: set your Anthropic API key to unlock structured
  narrative summaries for each production
- **Cover page**: customise the researcher photo, ORCID, LinkedIn link, and
  QR code
- **Package website**: <https://Paul1991-ux.github.io/publigraphics>
- **Source code**: <https://github.com/Paul1991-ux/publigraphics>
